---
title: "Agent Trust Analysis (Trust Exp. 2)"
author: "Madeline H. Pelgrim"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r echo=FALSE, message=FALSE, warning = FALSE}
#knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lmerTest)
library(rstatix)
library(emmeans)
library(plotrix)
library(cowplot)
```

# Import and Prepare Cleaned Data

*Data from qualtrics cleaned using dog_trust_cleaning.R file as in Experiment 1*

```{r}

analysis_df <- read_csv("trust_exp2_data.csv", show_col_types = FALSE)

##Create dataframe with summed scores (not broken down by MDMT V2 Subscore)
summed_df <- analysis_df %>%
  ungroup() %>%
  mutate(final_score = push_reliable_score + push_ethical_score + push_competent_score + push_benevolent_score + push_transparent_score, 
    baseline_score = reliable_score + ethical_score + competent_score + benevolent_score, transparent_score) %>%
  mutate(magnitude_change_score = abs(baseline_score - final_score), 
         change_score = final_score - baseline_score,
         social_scenario = ifelse(scenario_num == 3, 1,
                                  ifelse(scenario_num == 4, 2, scenario_num))) %>%
  mutate(social_scenario = replace(social_scenario, social_scenario == 5, 1)) %>%
  select(scenario, final_score, baseline_score, Agent, ParticipantID, scenario_num, Push_Direction, magnitude_change_score, change_score, social_scenario, scenario_order) %>%
  distinct()

```

#### Effects of social vs. less-social scenarios (collapsing across our 5)

```{r}
##Prepare dataframe
social_df <- summed_df %>%
  ungroup() %>%
  group_by(ParticipantID, social_scenario) %>%
  summarise(final_score = mean(final_score), baseline_score = mean(baseline_score),
            magnitude_change_score = mean(magnitude_change_score)) %>%
  group_by(ParticipantID) %>%
  filter(n() > 1) %>%
  ungroup()

##Gather summary statistics
social_df %>%
group_by(social_scenario) %>%
summarise(
mean_b = mean(baseline_score, na.rm = TRUE),
sd_b = sd(baseline_score, na.rm = TRUE),
mean_c = mean(magnitude_change_score, na.rm= TRUE),
sd_c = sd(magnitude_change_score, na.rm = TRUE)
)
        

##Are the baseline scores different between social and less social scenarios? 
t.test(baseline_score ~ social_scenario, data = social_df, paired = TRUE)


##Are the magnitude change scores different between social and less social scenarios? 
t.test(magnitude_change_score ~ social_scenario, data = social_df, paired = TRUE)

```

# Baseline Trust by Agent ANOVA

```{r}
##Baseline Scores

anova_df <- summed_df %>%
  group_by(ParticipantID) %>%
  summarise(Agent = Agent, baseline_score = mean(baseline_score)) %>%
  distinct()

anova_df %>%
  group_by(Agent) %>%
summarise(
  mean = mean(baseline_score, na.rm = TRUE),
  sd = sd(baseline_score, na.rm = TRUE))

agent_anova <- aov(baseline_score ~ Agent, data = anova_df)


summary(agent_anova)

TukeyHSD(agent_anova)
```

# Change by Agent and Direction

```{r}

##model without interaction
model0 <- lmer(magnitude_change_score ~ Push_Direction + Agent + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", Agent = "contr.sum"), data = summed_df)


##model with interaction
model1 <- lmer(magnitude_change_score ~ Push_Direction*Agent + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", Agent = "contr.sum"), data = summed_df)


anova(model0, model1) ##interaction does improve fit so proceed with interaction model

##summary of main effects
Anova(model0, type = 3)

summary(model0)
#follow-up on differences
pairs(emmeans(model0, ~ Agent))
```

### How many of the change scores match the valence of the push direction?

```{r}
summed_df %>%
  group_by(Push_Direction) %>%
summarise(
  mean = mean(change_score, na.rm = TRUE),
  sd = sd(change_score, na.rm = TRUE))

analysis_df %>%
  group_by(scenario_num, ParticipantID) %>%
  summarise(value = sum(value), Push_Direction = Push_Direction) %>%
  filter(!is.na(value)) %>%
  distinct() %>%
  mutate(sign = if_else(value > 0, "positive", "negative")) %>%
  group_by(Push_Direction, sign) %>%
  summarise(count = n(), .groups = "drop")
```

# Change by Agent and MDMT Subdimension

```{r}
model0.1 <- lmer(change_magnitude ~ Agent + measured_dimension + (1|ParticipantID), contrasts = list(Agent = "contr.sum", measured_dimension = "contr.sum"), data = analysis_df)

model1.1 <- lmer(change_magnitude ~ Agent*measured_dimension + (1|ParticipantID), contrasts = list(Agent = "contr.sum", measured_dimension = "contr.sum"), data = analysis_df)

anova(model0.1, model1.1) ###Interaction does improve model fit so proceed with

summary(model1.1)

Anova(model1.1, type = 3)

#follow-up on differences
pairs(emmeans(model1.1, ~ Agent | measured_dimension))

```

### Descriptives for the subdimension model

```{r}

analysis_df %>%
  group_by(Agent) %>%
summarise(
  mean = mean(change_magnitude, na.rm = TRUE),
  sd = sd(change_magnitude, na.rm = TRUE))


analysis_df %>%
  group_by(Push_Direction) %>%
summarise(
  mean = mean(change_magnitude, na.rm = TRUE),
  sd = sd(change_magnitude, na.rm = TRUE))
```

# Visualize differences by agent

### Baseline Trust

```{r}
agent_colors <- c(
  dog = "#A6CEE3",
  robot = "#1F78B4",
  person = "#B2DF8A"
)


analysis_df %>%
  select(ParticipantID, reliable_score:benevolent_score, Agent) %>%
  pivot_longer(!c(ParticipantID, Agent), 
               names_to = "measured_dimension", 
               values_to = "baseline_score") %>% 
  mutate(measured_dimension = str_remove(as.character(measured_dimension), "_score")) %>%
 group_by(measured_dimension, Agent) %>%
  summarise(mean = mean(baseline_score), 
            se = std.error(baseline_score)) %>%
  ungroup() %>%
  mutate(Agent = factor(Agent, levels = c("robot", "dog", "person")))  %>%
  ggplot(aes(x = measured_dimension, y = mean, fill = Agent)) + 
  geom_bar(stat = 'identity', position = "dodge") +
  theme(panel.background = element_blank(),
        axis.line = element_line(color = "black"),
         panel.grid.major = element_line(color = "grey85"),
        axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(face = "bold", size = 12)) + 
  scale_fill_manual(breaks=c("robot","dog","person"), values = agent_colors) +
  geom_errorbar(aes(ymin=mean - se, ymax=mean+se), width=.2,
                 position=position_dodge(.9)) + ylim(0,7) + xlab("Measured Dimension") + 
  ylab("Mean Baseline Score with Standard Error")

```

### Change in Trust

```{r}
#library("RColorBrewer")

#brewer.pal(3, "Paired")

agent_colors <- c(
  dog = "#A6CEE3",
  robot = "#1F78B4",
  person = "#B2DF8A"
)


up_graph <- analysis_df %>%
  filter(Push_Direction == "UP") %>%
  group_by(measured_dimension, Agent, Push_Direction) %>%
  summarise(mean = mean(value), 
            se = std.error(value)) %>%
  ungroup() %>%
  mutate(measured_dimension = as.character(measured_dimension)) %>%
  mutate(measured_dimension = replace(measured_dimension, measured_dimension == "B", "Benevolent"), 
         measured_dimension = replace(measured_dimension, measured_dimension == "E", "Ethical"),
         measured_dimension = replace(measured_dimension, measured_dimension == "R", "Reliable"),
         measured_dimension = replace(measured_dimension, measured_dimension == "C", "Competent"),
         measured_dimension = replace(measured_dimension, measured_dimension == "T", "Transparent")) %>%
  mutate(Agent = factor(Agent, levels = c("robot", "dog", "person")))  %>%
  ggplot(aes(x = measured_dimension, y = mean, fill = Agent)) + 
  geom_bar(stat = 'identity', position = "dodge") + ylim(0, 2.5) +
  theme(legend.position = "none", 
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
         panel.grid.major = element_line(color = "grey85")) + 
  scale_fill_manual(values = agent_colors) + 
  geom_errorbar(aes(ymin=mean - se, ymax=mean+se), width=.2,
                 position=position_dodge(.9))
  #scale_fill_brewer(palette = "Paired") 
 
down_graph <- analysis_df %>%
  filter(Push_Direction == "DOWN") %>%
  group_by(measured_dimension, Agent, Push_Direction) %>%
  mutate(Agent = factor(Agent, levels = c("robot", "dog", "person"))) %>%
  summarise(mean = mean(value),
            se = std.error(value)) %>%
  ggplot(aes(x = measured_dimension, y = mean, fill = Agent)) + 
  geom_bar(stat = 'identity', position = "dodge") + ylim(-2.5, 0) + 
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey85")) + 
  scale_x_discrete(position = "top") + 
  scale_fill_manual(breaks=c("robot","dog","person"), values = agent_colors) + 
  geom_errorbar(aes(ymin=mean - se, ymax=mean+se), width=.2,
                 position=position_dodge(.9))
  #scale_fill_brewer(palette = "Paired")


legend <- get_legend(down_graph)
  
down_graph_a <- down_graph + theme(legend.position = "none")


# Stack the graphs vertically
combined_graphs <- plot_grid(
  up_graph,
  down_graph_a,
  ncol = 1,
  align = "v"
)

# Add the legend to the right
stack_with_legend <- plot_grid(
  combined_graphs,
  legend,
  ncol = 2,
  rel_widths = c(3, 1),
  align = "h"
)


final_plot <- ggdraw() +
  draw_label("Average Change Score with Standard Error", x = 0.02, y = 0.5, angle = 90,
             hjust = 0.5, fontface = "bold", size = 12) +
  draw_label("Measured Dimension", x = 0.4, y = 0.03, angle = 0,
             hjust = 0.5, fontface = "bold", size = 12) +
  draw_plot(stack_with_legend, x = 0.05, y = 0.05, width = 0.9, height = 0.9)

# Print the final plot
print(final_plot)


```

# Participant Factors

### Presentation Order

```{r}
##Change Scores
anova_test(data = summed_df, dv = change_score, wid = ParticipantID, within = scenario_order)

##Baseline Scores
anova_test(data = summed_df, dv = baseline_score, wid = ParticipantID, within = scenario_order)

```

### Dog & Robot Factors

##### Run correlations on participants who saw dog or robot agents. First prepare the data

```{r}

analysis_df2 <- analysis_df %>%
  select(ParticipantID, Experience, Knowledge)


factors <- merge(summed_df, analysis_df2)

for_cors <- factors %>%
  filter(Agent != "person") %>%
  group_by(ParticipantID) %>%
  summarise(Experience = Experience, 
            Agent = Agent,
            Knowledge = Knowledge,
            baseline_score = mean(baseline_score), 
            magnitude_change_score = mean(magnitude_change_score), 
            change_score = mean(change_score), 
            final_score = mean(final_score)) %>%
  unique()
```

### Do knowledge and experience with dogs predict trust??

##### Experience with dogs

```{r}
dog_cors <- for_cors %>% filter(Agent == "dog")


table(dog_cors$Experience)


cor.test(dog_cors$Experience, dog_cors$baseline_score, method = "spearman")


cor.test(dog_cors$Experience, dog_cors$magnitude_change_score, method = "spearman")
```

##### Knowledge about dogs

```{r}

table(dog_cors$Knowledge)


cor.test(dog_cors$Knowledge, dog_cors$baseline_score, method = "spearman")


cor.test(dog_cors$Knowledge, dog_cors$magnitude_change_score, method = "spearman")
```

### Do knowledge and experience with robots predict trust??

##### Experience with robots

```{r}
robot_cors <- for_cors %>% filter(Agent == "robot")


table(robot_cors$Experience)


cor.test(robot_cors$Experience, robot_cors$baseline_score, method = "spearman")


cor.test(robot_cors$Experience, robot_cors$magnitude_change_score, method = "spearman")
```

##### Knowledge about robots

```{r}

table(robot_cors$Knowledge)


cor.test(robot_cors$Knowledge, robot_cors$baseline_score, method = "spearman")


cor.test(robot_cors$Knowledge, robot_cors$magnitude_change_score, method = "spearman")
```
