---
title: "Dog Trust Analysis"
author: "Madeline H. Pelgrim"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r echo=FALSE, message=FALSE, warning = FALSE}
library(tidyverse)
library(lmerTest)
library(rstatix)
library(emmeans)
library(plotrix)
```

# Import and Prepare Cleaned Data

*Data from qualtrics cleaned using dog_trust_cleaning.R file*

```{r}

analysis_df <- read_csv("trust_exp1_data.csv")

##Create dataframe with summed scores (not broken down by MDMT V1 Subscore)
summed_df <- analysis_df %>%
  ungroup() %>%
  mutate(final_score = push_reliable_score + push_ethical_score + push_capable_score + push_sincere_score, 
    baseline_score = reliable_score + ethical_score + capable_score + sincere_score) %>%
  mutate(magnitude_change_score = abs(baseline_score - final_score), 
         change_score = final_score - baseline_score,
         social_scenario = ifelse(scenario_num == 3, 1,
                                  ifelse(scenario_num == 4, 2, scenario_num))) %>%
  select(scenario, final_score, baseline_score, ParticipantID, scenario_num, Push_Direction, magnitude_change_score, change_score, social_scenario, original_sample, scenario_order) %>%
  distinct()
```

## Are MDMT scores different across our scenarios?

#### Summed baseline trust scores as a function of the scenario

```{r}

##Baseline Scores
anova_test(data = summed_df, dv = baseline_score, wid = ParticipantID, within = scenario_num)

##Magnitude Change Scores
anova_test(data = summed_df, dv = magnitude_change_score, wid = ParticipantID, within = scenario_num)
```

#### Effects of social vs. less-social scores (collapsing across our 4)

```{r}

##Prepare dataframe
social_df <- summed_df %>%
  ungroup() %>%
  group_by(ParticipantID, social_scenario) %>%
  summarise(final_score = mean(final_score), baseline_score = mean(baseline_score),
            magnitude_change_score = mean(magnitude_change_score)) %>%
  group_by(ParticipantID) %>%
  filter(n() > 1) %>%
  ungroup()

##Gather summary statistics
social_df %>%
group_by(social_scenario) %>%
summarise(
mean_b = mean(baseline_score, na.rm = TRUE),
sd_b = sd(baseline_score, na.rm = TRUE),
mean_c = mean(magnitude_change_score, na.rm= TRUE),
sd_c = sd(magnitude_change_score, na.rm = TRUE)
)
        

##Are the baseline scores different between social and less social scenarios? 
t.test(baseline_score ~ social_scenario, data = social_df, paired = TRUE)


##Are the magnitude change scores different between social and less social scenarios? 
t.test(magnitude_change_score ~ social_scenario, data = social_df, paired = TRUE)

```

# Primary Model - Change in Trust Summed

```{r}
##model without interaction
model0 <- lmer(change_score ~ Push_Direction + scenario + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", scenario = "contr.sum"), data = summed_df)

##model with interaction
model1 <- lmer(change_score ~ Push_Direction*scenario + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", scenario = "contr.sum"), data = summed_df)

anova(model0, model1) ##interaction does NOT improve fit so don't proceed with this model

##summary of main effects
Anova(model0, type = 3)

summary(model0)

#follow-up on scenario differences
pairs(emmeans(model0, ~ scenario))
```

### How many of the change scores match the valence of the push direction?

```{r}
analysis_df %>%
  group_by(scenario_num, ParticipantID) %>%
  summarise(value = sum(value), Push_Direction = Push_Direction) %>%
  filter(!is.na(value)) %>%
  distinct() %>%
  mutate(sign = if_else(value > 0, "positive", "negative")) %>%
  group_by(Push_Direction, sign) %>%
  summarise(count = n(), .groups = "drop")
```

### Visualization of change in trust scores

```{r}
summed_df %>%
  mutate(scenario = replace(scenario, scenario == "C", "Capable"), 
         scenario = replace(scenario, scenario == "E", "Ethical"),
         scenario = replace(scenario, scenario == "R", "Reliable"),
         scenario = replace(scenario, scenario == "S", "Sincere")) %>%
  ggplot(aes(x = scenario, y = change_score, fill = Push_Direction)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge") + ylab("Average Change in Trust") + xlab("MDMT V1 Pushed Dimension") + theme_minimal()
```

# Secondary Model - Change in Trust by Subdimension

```{r}
##model without interaction
model0.1 <- lmer(value ~ Push_Direction + pushed_dimension + measured_dimension + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", pushed_dimension = "contr.sum", measured_dimension = "contr.sum"), data = analysis_df)

##model with interaction
model1.1 <- lmer(value ~ Push_Direction*pushed_dimension + measured_dimension + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", pushed_dimension = "contr.sum", measured_dimension = "contr.sum"), data = analysis_df)

anova(model0.1, model1.1) ##interaction does NOT improve fit so don't proceed

##main effects
Anova(model0.1, type = 3)

summary(model0.1)

#follow-up on scenario differences
pairs(emmeans(model0.1, ~ measured_dimension))
```

# Comparison to Ullman & Malle, (2019)

### Prepare the data

```{r}
##Read in data from Ullman & Malle (2019)
robot_trust <- read_csv("ullman_malle_2019.csv") %>%
  select(ParticipantID = ResponseId, 
         reliable_score = MDMT_Rel, 
         ethical_score = MDMT_Eth, 
         capable_score = MDMT_Comp,
         sincere_score = MDMT_Sinc) %>%
  mutate(species = "robot")

##Prepare for the data to be combined
species_comparison <- analysis_df %>%
  mutate(ParticipantID = as.character(ParticipantID)) %>%
  group_by(ParticipantID) %>%
  summarise(across(c(reliable_score, 
                     ethical_score, 
                     capable_score, 
                     sincere_score), mean, na.rm = TRUE)) %>%
  mutate(species = "dog") %>%
  bind_rows(robot_trust) %>%
  pivot_longer(cols = c(reliable_score, ethical_score, capable_score, sincere_score),
               names_to = "dimension",
               values_to = "value") %>%
  mutate(dimension = as.factor(dimension), 
         species = as.factor(species)) %>%
  filter(!is.na(value))

```

### Gather summary statistics

```{r}
##MDMT score by agent type only (collapsed across dimensions)
species_comparison %>%
group_by(species) %>%
  summarise(mean_score = mean(value, na.rm = TRUE), 
            sd_score = sd(value, na.rm = TRUE))

##MDMT score by dimension only (collapsed across agents)
species_comparison %>%
group_by(dimension) %>%
  summarise(mean_score = mean(value, na.rm = TRUE), 
            sd_score = sd(value, na.rm = TRUE))

##MDMT score broken down by both agent and dimension
species_comparison %>%
group_by(species, dimension) %>%
  summarise(mean_score = mean(value, na.rm = TRUE), 
            sd_score = sd(value, na.rm = TRUE))

```

### Run the 2 (agent) by 4(dimension) ANOVA

```{r}
summary(aov(value ~ dimension + species, data = species_comparison))

```

### Visualize differences between trust in dogs and robots on MDMT V1

```{r}
species_comparison %>%
  mutate(dimension = str_remove(as.character(dimension), "_score")) %>%
  group_by(species, dimension) %>%
  summarise(avg = mean(value), 
            se = std.error(value)) %>%
  ggplot(aes(x = dimension, y = avg, fill = species)) + 
  geom_bar(stat = "identity", position = "dodge") + ylab("Average Baseline Trust with Standard Error") + xlab("MDMT V1 Measured Dimension") + theme_minimal() + ylim(0, 7) + geom_errorbar(aes(ymin=avg - se, ymax=avg+se), width=.2,
                 position=position_dodge(.9)) + labs(fill = "Agent") + scale_fill_brewer(palette = "Paired")
```

# Participant Factors

### Prepare the data

```{r}

analysis_df2 <- analysis_df %>%
  select(ParticipantID, PAS_sum, Experience_Rate, Knowledgeable)


dog_priors <- merge(summed_df, analysis_df2)

baseline <- dog_priors %>%
  group_by(ParticipantID) %>%
  summarise(PAS = PAS_sum, Experience_Rate = Experience_Rate, 
            Knowledgeable = Knowledgeable,
            baseline_score = mean(baseline_score), 
            magnitude_change_score = mean(magnitude_change_score), 
            change_score = mean(change_score), 
            final_score = mean(final_score)) %>%
  unique()

```

### Impact of Pet Attitude Scale

```{r}

cor.test(baseline$PAS, baseline$baseline_score)

cor.test(baseline$PAS, baseline$magnitude_change_score)

```

### Impact of experience with dogs

```{r}

##descriptives
table(baseline$Experience_Rate)

##Correlation - baseline
cor.test(baseline$Experience_Rate, baseline$baseline_score, method = "spearman")

##Correlation - magnitude change
cor.test(baseline$Experience_Rate, baseline$magnitude_change_score, method = "spearman")

```

### Impact of knowledge about dogs

```{r}

##descriptives
table(baseline$Knowledgeable)

##Correlation - baseline
cor.test(baseline$Knowledgeable, baseline$baseline_score, method = "spearman")

##Correlation - magnitude change
cor.test(baseline$Knowledgeable, baseline$magnitude_change_score, method = "spearman")

```

### Impact of Presentation Order

```{r}

##Change Scores
anova_test(data = summed_df, dv = change_score, wid = ParticipantID, within = scenario_order)

##Baseline Scores
anova_test(data = summed_df, dv = baseline_score, wid = ParticipantID, within = scenario_order)

```

# Supplemental - complete analyses with only the original sample

### Prepare the data

```{r}
##Create the smaller sample
subset_df <- analysis_df %>%
  filter(original_sample == "YES") 

##Create the summed dataset with the smaller sample
summed_subset_df <- subset_df %>%
  ungroup() %>%
  mutate(final_score = push_reliable_score + push_ethical_score + push_capable_score + push_sincere_score, 
    baseline_score = reliable_score + ethical_score + capable_score + sincere_score) %>%
  mutate(magnitude_change_score = abs(baseline_score - final_score), 
         change_score = final_score - baseline_score,
         social_scenario = ifelse(scenario_num == 3, 1,
                                  ifelse(scenario_num == 4, 2, scenario_num))) %>%
  select(scenario, final_score, baseline_score, ParticipantID, scenario_num, Push_Direction, magnitude_change_score, change_score, social_scenario, original_sample, scenario_order) %>%
  filter(!is.na(scenario_num)) %>%
  group_by(ParticipantID) %>%
  filter(n() > 0) %>%
  ungroup() %>%
  distinct()
```

### Original Sample - are MDMT scores different across our scenarios?

```{r}
##Baseline Scores
anova_test(data = summed_subset_df, dv = baseline_score, wid = ParticipantID, within = scenario_num)

##Magnitude Change Scores
anova_test(data = summed_subset_df, dv = magnitude_change_score, wid = ParticipantID, within = scenario_num)
```

### Original Sample - Effects of social vs. less-social scores

```{r}
##Prepare dataframe
subset_social_df <- summed_subset_df %>%
  ungroup() %>%
  group_by(ParticipantID, social_scenario) %>%
  summarise(final_score = mean(final_score), baseline_score = mean(baseline_score),
            magnitude_change_score = mean(magnitude_change_score)) %>%
  group_by(ParticipantID) %>%
  filter(n() > 1) %>%
  ungroup()

##Gather summary statistics
subset_social_df %>%
  group_by(social_scenario) %>%
  summarise(
    mean_b = mean(baseline_score, na.rm = TRUE),
    sd_b = sd(baseline_score, na.rm = TRUE),
    mean_c = mean(magnitude_change_score, na.rm= TRUE),
    sd_c = sd(magnitude_change_score, na.rm = TRUE))

##Are the baseline scores different between social and less social scenarios?
t.test(baseline_score ~ social_scenario, data = subset_social_df, paired = TRUE)

##Are the magnitude change scores different between social and less social scenarios?
t.test(magnitude_change_score ~ social_scenario, data = subset_social_df, paired = TRUE)
```

### Original Sample Primary Model - Change in Trust Summed

```{r}
##run identical model to main manuscript
model0 <- lmer(change_score ~ Push_Direction + scenario + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", scenario = "contr.sum"), data = summed_subset_df)

Anova(model0, type = 3)

summary(model0)

#follow-up on scenario differences
pairs(emmeans(model0, ~ scenario))
```

### Original Sample - how many change scores match the valence of the push direction?

```{r}
subset_df %>%
  group_by(scenario_num, ParticipantID) %>%
  summarise(value = sum(value), Push_Direction = Push_Direction) %>%
  filter(!is.na(value)) %>%
  distinct() %>%
  mutate(sign = if_else(value > 0, "positive", "negative")) %>%
  group_by(Push_Direction, sign) %>%
  summarise(count = n(), .groups = "drop")
```

### Original Sample - descriptives for change in trust

```{r}
summed_subset_df %>%
  group_by(Push_Direction) %>%
  summarise(
    mean_change = mean(change_score, na.rm = TRUE),
    sd_change = sd(change_score, na.rm = TRUE))

```

### Original Sample - Comparisons to Ullman & Malle (2019)

```{r}
##Prepare for the data to be combined
subset_species_comparison <- analysis_df %>%
  filter(original_sample == "YES") %>%
  mutate(ParticipantID = as.character(ParticipantID)) %>%
  group_by(ParticipantID) %>%
  summarise(across(c(reliable_score, 
                     ethical_score, 
                     capable_score, 
                     sincere_score), mean, na.rm = TRUE)) %>%
  mutate(species = "dog") %>%
  bind_rows(robot_trust) %>%
  pivot_longer(cols = c(reliable_score, ethical_score, capable_score, sincere_score),
               names_to = "dimension",
               values_to = "value") %>%
  mutate(dimension = as.factor(dimension), 
         species = as.factor(species)) %>%
  filter(!is.na(value))

##MDMT score by agent type only (collapsed across dimensions)
subset_species_comparison %>%
group_by(species) %>%
  summarise(mean_score = mean(value, na.rm = TRUE), 
            sd_score = sd(value, na.rm = TRUE))


##MDMT score by dimension only (collapsed across agents)
subset_species_comparison %>%
group_by(dimension) %>%
  summarise(mean_score = mean(value, na.rm = TRUE), 
            sd_score = sd(value, na.rm = TRUE))

##MDMT score broken down by both agent and dimension
subset_species_comparison %>%
group_by(species, dimension) %>%
  summarise(mean_score = mean(value, na.rm = TRUE), 
            sd_score = sd(value, na.rm = TRUE))

```

### **Original Sample - Run the 2 (agent) by 4(dimension) ANOVA**

```{r}
summary(aov(value ~ dimension + species, data = subset_species_comparison))

```

### Original Sample - Secondary Model

```{r}

model0.1 <- lmer(value ~ Push_Direction + pushed_dimension + measured_dimension + (1|ParticipantID), contrasts = list(Push_Direction = "contr.sum", pushed_dimension = "contr.sum", measured_dimension = "contr.sum"), data = subset_df)

Anova(model0.1, type = 3)

summary(model0.1)

#follow-up on scenario differences
pairs(emmeans(model0.1, ~ measured_dimension))

pairs(emmeans(model0.1, ~ pushed_dimension))

```

### Original Sample Participant Factors

#### Prepare the data

```{r}

analysis_df2 <- subset_df %>%
  select(ParticipantID, PAS_sum, Experience_Rate, Knowledgeable)

dog_priors <- merge(summed_subset_df, analysis_df2)

baseline <- dog_priors %>%
  group_by(ParticipantID) %>%
  summarise(PAS = PAS_sum, Experience_Rate = Experience_Rate, 
            Knowledgeable = Knowledgeable,
            baseline_score = mean(baseline_score), 
            magnitude_change_score = mean(magnitude_change_score), 
            change_score = mean(change_score), 
            final_score = mean(final_score)) %>%
  unique() %>%
  mutate(Experience_Rate = as.numeric(Experience_Rate), 
         Knowledgeable = as.numeric(Knowledgeable))

```

#### Impact of Pet Attitude Scale

```{r}
cor.test(baseline$PAS, baseline$baseline_score)

cor.test(baseline$PAS, baseline$magnitude_change_score)

```

#### Impact of experience with dogs

```{r}
##descriptives
table(baseline$Experience_Rate)

##Correlation - baseline
cor.test(baseline$Experience_Rate, baseline$baseline_score, method = "spearman")

##Correlation - magnitude change score
cor.test(baseline$Experience_Rate, baseline$magnitude_change_score, method = "spearman")

```

#### Impact of knowledge about dogs

```{r}

##descriptives
table(baseline$Knowledgeable)

##Correlation - baseline
cor.test(baseline$Knowledgeable, baseline$baseline_score, method = "spearman")

##Correlation - magnitude change score
cor.test(baseline$Knowledgeable, baseline$magnitude_change_score, method = "spearman")

```

### Impact of Presentation Order

```{r}

##Change Scores
anova_test(data = summed_subset_df, dv = change_score, wid = ParticipantID, within = scenario_order)

##Baseline Scores
anova_test(data = summed_subset_df, dv = baseline_score, wid = ParticipantID, within = scenario_order)

```
